### 2021-05-03 article/article.tex

\documentclass[a4paper,article,14pt]{extarticle}

\usepackage{allpackages}

\begin{document}
\newgeometry{left=30mm, top=20mm, right=15mm, bottom=20mm, nohead}

\input{titlepage.tex}

\tableofcontents
\newpage

\input{intro.tex}
\newpage

\input{thesis.tex}
\newpage

\input{literature_observation.tex}
\newpage

\input{theory.tex}
\newpage

\input{solution.tex}
\newpage

\input{results.tex}
\newpage

\input{findings.tex}
\newpage

\input{conclusion.tex}
\newpage

\input{bibliography.tex}
\newpage

\input{appendix.tex}
\restoregeometry
\end{document}