### 2021-05-03 article/bibliography.tex

\begin{thebibliography}{1}
\bibitem{Audrey}
Davis K. N., Biddulph R., Balashek S. Automatic recognition of spoken digits // The Journal of the Acoustical Society of America, 1952. Vol. 24, No 6. P. 637-642

\bibitem{Harpy}
Newell A. Harpy, production systems and human cognition // Research Showcase @ Carnegie Mellon University, 1978

\bibitem{spectra}
Plomp R., Pols L. C. W., van der Geer J.P. Dimensional analysis of vowel spectra // The Journal of the Acoustical Society of America, 1967. Vol. 41, P. 707-712

\bibitem{MFCC}
Davis S., Mermelstein P. Comparison of parametric representations for monosyllabic word recognition in continuously spoken sentences // IEEE Transactions on Acoustics, Speech, and Signal Processing , 1980. Vol. ASSP-28, No. 4, P. 357-366

\bibitem{SignalPreprocessing}
Rabiner L.R., Sambur, M.R. An Algorithm for Determining the Endpoints of Isolated Utterances // The Bell System Technical Journal 1975 Vol. 54, No. 2, P. 297-315

\bibitem{Perceptron}
Rosenblatt, Frank. Principles of Neurodynamics: Perceptrons and the Theory of Brain Mechanisms. Spartan Books, Washington DC, 1961

\bibitem{CNN}
LeCun Y., Bengio Y.. Convolutional networks for images, speech, and time-series // The Handbook of Brain Theory and Neural Networks, 1995, MIT 
\end{thebibliography}